Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer

### 2020-03-23 manuscript/manuscript.tex
went over text up to 3.5

@@

\begin{abstract}
Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building. It does so by jointly considering the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network. In this work, we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype. We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women. We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods. By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility. Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$ ) and occupy more central positions in the network than average. Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer. Finally, we find a significantly large overlap between the genes in the consensus network and the genes significantly associated in the largest GWAS on susceptibility to breast cancer. In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
\end{abstract}

@@

In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \cite{bush_chapter_2012}. To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering the disease of interest, and the controls from the general population. Then, a per-SNP statistical association test is conducted (e.g. logistic regression). Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort. Once the risk SNPs have been discovered, they can be used for risk assessment, and to deepen our understanding of the disease.

GWAS have successfully identified thousands of variants underlying many common diseases \cite{buniello_nhgri-ebi_2019}. However, this experimental setting also presents intrinsic challenges. Some of them stem from the high-dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped. This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}. And it is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}. Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the aforementioned Bonferroni. However, Bonferroni is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}. Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood. On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}. The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism. Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.

In this regard many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}. As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them. These relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction. Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}. In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}. This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes. Additionally, as genes that contribute to a disease tend to participate in similar biological functions, studying the neighborhood of known disease genes is effective at identifying new ones \cite{huang_systematic_2018}.

Network-based, biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}. In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge. Then, the problem becomes finding a functionally-related set of highly-scoring genes. Different solutions have been proposed to this problem, often stemming from divergent mathematical frameworks and considerations of what the optimal solution looks like. Some methods strongly constrain the problem to certain kinds of subnetworks. Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e. instances were both a gene and its direct interactors are associated with the disease. Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnected genes with high association scores. However, they differ in their tolerance to the inclusion of low scoring nodes, and the possible number of disconnected subnetworks in the solution. Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.

@@

We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy - Weinberg equilibrium in controls (P-value \textless~0.001), and those missing on more than 10\% of the samples. A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed. In addition, we removed the samples with more than 10\% missing genotypes. After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, and 3 additional controls having a high relatedness score). Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis). The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.

@@

\subsubsection{Mathematical notation}
\label{methods:notation}
In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$. $V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$. Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges. When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$. A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.

On top of a weight, nodes have other properties, provided by the topology of the graph. In this article we focus on two: degree centrality, and betweenness centrality. The degree centrality, or degree, is the number of edges that a node has. The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.

@@

\paragraph{Gene network}
Out of the six methods tested, five use a gene-gene interaction network (Section~\ref{methods:methods}). Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), we used protein-protein interaction networks (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it. We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}. Unless specified otherwise, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network. Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols. The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).

@@

@@

\subsubsection{Studied methods}
\label{methods:methods}

\begin{table}[htbp]
  \caption{\label{tab:method_comparison} Summary of the differences between the studied algorithms.}
  \centering
  \begin{threeparttable}
    \begin{tabular}{l|llllllll}
      Method & Field & Nodes & Exhaustive & Solution & Comp. & Input & Scoring & Ref.\\
      \hline
      dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
      heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
      HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
      LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
      SigMod & GWAS & Genes & Yes & Module & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
    \end{tabular}
    \begin{tablenotes}
      \footnotesize{
        \item \textbf{\emph{Field}}: field in which the algorithm was developed. \textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network. \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored. \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected. \textbf{\emph{Comp.}}: number of connected subnetworks in the solution. \textbf{\emph{Input}}: genotype data or GWAS summary statistics. \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
      }
    \end{tablenotes}
  \end{threeparttable}
\end{table}

Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.). But accounting for that is not the only choice a developer of a network method needs to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc. Unsurprisingly, multiple solutions have been proposed. We examined six of them: five that explore the PPIN, and one which explores SNP networks. We selected methods that were open source, had an implementation available, and an accessible documentation. Their main differences are summarized in Table~\ref{tab:method_comparison}.

@@

\begin{enumerate}
\item Select a seed node.
\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph $S$ as

\begin{equation*} 
Z_m = \frac{1}{\sqrt{k}} \sum_{i \in S} z_i
\end{equation*}

where \emph{k} is the number of genes in the subgraph; z\textsubscript{i} is the Z score of gene $i$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
\item Identify neighboring nodes of $S$ i.e. nodes at distance \(\le\) d. We set d~=~2.
\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r). In our experiments, we set r~=~0.1.
\item Repeat 2-4 until no further enlargement is possible.
\end{enumerate}

Lastly, the module's Z-score is normalized as

@@

We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}. We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
\end{description}

\begin{description}
\item[{heinz}] The goal of heinz is to identify the highest-scored connected subgraph on the network \cite{dittrich_identifying_2008}. The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive when there is. This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired FDR. Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitely long computational time. To solve it efficiently it is re-casted as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S):

@@

were p(v)~=~w(v) - w' is the \emph{profit} of adding a node, c(e)~=~w' is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$. All three are positive quantities. Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, although in practice is often fast and optimal, neither is guaranteed. We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.

\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}. To that end, it considers both the local topology of the network and the scores of the nodes. The former is captured by an insulated heat diffusion process: at the beginning, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}). This process continues until equilibrium is reached, and results in a similarity matrix $F$. $F$ is used to compute the similarity matrix $E$ that accounts also for similarities in node scores as

\begin{equation*} 
E = F \operatorname{diag}(w(V)), 
\end{equation*}

where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal. We scored the nodes as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be. In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}. To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$. Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores. HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\). Both parameters are set automatically by the algorithm, and are robust \cite{leiserson_pan-cancer_2015}. HotNet2 is implemented in Python \cite{hotnet2}.

@@

\begin{equation}
\label{eq:scones}
\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
\end{equation}

where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model. Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure. As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds. In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI. We used the version on SConES implemented in the R package martini \cite{martini}.
\end{description}

\begin{description}
\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}. It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the adjacency matrix rather than the Laplacian matrix (Section~\ref{methods:notation}), to quantify solutions containing many edges.

>>>>

\usepackage{lmodern} % Chloe: better font rendering

% Chloe's commands
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem} 
\newcommand{\cazcom}[2]{{\uline{#1}}\unskip\space\textcolor{MidnightBlue}{[CA: #2]}}
\newcommand{\caz}[2]{{\sout{#1}}\unskip\space\textcolor{MidnightBlue}{#2}}
%%% To get the original text before my comments:
% \newcommand{\cazcom}[2]{#1}
% \newcommand{\caz}[2]{#1}
%%% To get the corrected text and remove my comments:
% \newcommand{\cazcom}[2]{#1}
% \newcommand{\caz}[2]{#2}

@@

\begin{abstract}
Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building. \caz{It does so by jointly considering}{Indeed, it allows to jointly consider} the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network. In this work, we study the utility of six \caz{network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype}{methods that identify subnetworks with higher association scores to a phenotype to discover new biomarkers for breast cancer susceptibility}. We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women. We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods. By trading statistical \caz{astringency}{stringency} for biological meaningfulness, most network methods \caz{get}{give} more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility. Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$) and occupying more central positions in the network than average. Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer. Finally, we find a significantly large overlap between the genes in the consensus network and the genes significantly associated in the largest GWAS on susceptibility to breast cancer. In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
\cazcom{}{Add a sentence to the effect that the computational tools are available on github for ease of application to other data sets?}
\end{abstract}

@@

In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \cite{bush_chapter_2012}. To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering the disease of interest, and the controls from the general population. Then, a per-SNP statistical association test is conducted (e.g. \cazcom{\caz{}{based on }logistic regression}{You're doing a chi2 allelic test later on, so maybe use this here rather than talking about logistic regression?}). Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort. Once the risk SNPs have been discovered, they can be used for risk assessment, and to deepen our understanding of the disease.

GWAS have successfully identified thousands of variants underlying many common diseases \cite{buniello_nhgri-ebi_2019}. However, this experimental setting also presents intrinsic challenges. Some of them stem from the high dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped. This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}. \caz{And it}{This} is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}. Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the \caz{aforementioned}{previously mentioned} Bonferroni\caz{}{correction}. However, Bonferroni \caz{}{correction} is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}. Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood. On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}. The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism. Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.

\caz{In this regard}{For this reason,} many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}. As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them. These relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction. Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}. In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}. This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes. Additionally, as genes that contribute to a disease tend to participate in similar biological functions, studying the neighborhood of known disease genes is effective at identifying new ones \cite{huang_systematic_2018}.

Network-based biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}. In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge. Then, the problem becomes finding a functionally-related set of highly-scoring genes. Different solutions have been proposed to this problem, often stemming from divergent mathematical frameworks and considerations of what the optimal solution looks like. Some methods strongly constrain the problem to certain kinds of subnetworks. Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e. instances were both a gene and its direct interactors are associated with the disease. Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnecting genes with high association scores. However, they differ in their tolerance to the inclusion of low-scoring nodes, and the possible number of \cazcom{\caz{disconnected subnetworks}{connected components}}{Are you worried people don't understand ``connected components''? If so, we should define it and use it.} in the solution. Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.

@@

We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P-value \textless~0.001), and those missing on more than 10\% of the samples. A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed. In addition, we removed the samples with more than 10\% missing genotypes. After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, and 3 additional controls having a high relatedness score). Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis). The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.

@@

\subsubsection{Mathematical notations}
\label{methods:notation}
In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$. $V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$. Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges. When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$. A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.

\caz{On top of}{In addition to} a weight, nodes have other properties, provided by the topology of the graph. In this article we focus on two\caz{}{of those}: degree centrality, and betweenness centrality. The degree centrality, or degree, is the number of edges that a node has. The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.

@@

\paragraph{Gene network}
Out of the six methods tested, \cazcom{five use a gene-gene interaction network}{SConES GI does too...} (Section~\ref{methods:methods}). Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), we used protein-protein interaction networks (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to them. We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}. Unless \caz{specified otherwise}{otherwise specified}, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network. Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols. The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).

@@

\cazcom{}{How many genes are involved in those interactions? In the ``SNP and gene association'' section, you describe 164\,037 SNPs mapping to 32\,767 gene symbols. How many of these genes and SNPs are involved in the aforementioned 142\,541 (resp. 173\,797) interactions? And what happens to the genes to which you have mapped SNPs but that are not involved in any HINT association: are they discarded, or do they remain as isolated nodes?}

@@

\cazcom{}{What are the sizes of the resulting GS, GM and GI networks? 164\,037 SNPs/nodes (or fewer if some aren't used on the gene-gene networks and you've removed them entirely from the analysis)? How many edges in the different networks? If you're worried about space, the sizes of the gene-gene and SNP-SNP networks from the previous paragraph and this one could be given in a table in the supplementaries.}

\subsubsection{Studied methods}
\label{methods:methods}
\cazcom{}{Maybe we could change ``studied'' in this section (title, text, tables...) to something that makes it sound a bit less like a school assignment, and also less like we decided to study those but could have studied others? Something like ``Network-based methods'' would have a bit more agency to it? Sorry if I'm nitpicking.}

\begin{table}[htbp]
  \caption{\label{tab:method_comparison} Summary of the differences between the studied algorithms.}
  \centering
  \begin{threeparttable}
    \begin{tabular}{l|llllllll}
      Method & Field & Nodes & Exhaustive & Solution & Comp. & Input & Scoring & Ref.\\
      \hline
      dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
      heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
      HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
      LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
      SigMod & GWAS & Genes & Yes & Module & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
    \end{tabular}
    \begin{tablenotes}
      \footnotesize{
        \item \textbf{\emph{Field}}: field in which the algorithm was developed. \textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network. \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored. \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected. \textbf{\emph{Comp.}}: number of connected \caz{subnetworks}{components} in the solution. \textbf{\emph{Input}}: genotype data or GWAS summary statistics. \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
      }
    \end{tablenotes}
  \end{threeparttable}
\end{table}

Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.). But \caz{accounting for that is not the only choice a developer of a network method needs to make}{this is not the only aspect to model when developing a network method}: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc. Unsurprisingly, multiple solutions have been proposed. We examined six of them: five that explore the PPIN, and one which explores SNP networks. We selected methods that were open source, had an implementation available, and an accessible documentation. Their main differences are summarized in Table~\ref{tab:method_comparison}.

@@

\begin{enumerate}
\item Select a seed node \caz{}{$i$ and form $S_i=\{i\}$}.
\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph \caz{$S$}{$S_i$} as

\begin{equation*} 
Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j
\end{equation*}

where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
\item Identify neighboring nodes of \caz{$S$}{$S_i$},  i.e. nodes at distance \(\le\) d. \caz{We set d~=~2.}{}
\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r). \caz{In our experiments, we set r~=~0.1.}{}
\item Repeat 2-4 until no further enlargement is possible.
\caz{}{\item Add $S_i$ to the list of subnetworks to return.}
\end{enumerate}
\cazcom{}{Missing here: how to select seed nodes. Is every node a seed node? Every node that has not yet been included in a subgraph?}

Lastly, \caz{the module's}{a subnetwork's} Z-score is normalized as

@@

We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}. \caz{}{We used the suggested hyperparameters $d = 2$ and $r = 0.1$} We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
\end{description}

\begin{description}
\item[{heinz}] The goal of heinz is to identify the highest-scored connected subgraph on the network \cite{dittrich_identifying_2008}. The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive when there is. This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired \caz{FDR}{false discovery rate (FDR)}. Thus formulated, the problem is NP-complete, and hence solving it would require a \caz{prohibitely}{prohibitively} long computational time. To solve it efficiently it is re-cast\caz{ed}{} as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:

@@

were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, \caz{$c(e)~=~w'$}{$c(e)=-w'$} is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ \caz{}{is the smallest node weight of $G$}. All three are positive quantities. Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, \caz{although} in practice is often fast and optimal, \caz{}{although} neither is guaranteed. We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.

\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}. To that end, it considers both the local topology of the network and the scores of the nodes. The former is captured by an insulated heat diffusion process: at \caz{the beginning}{initialization}, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}). This process continues until equilibrium is reached, and results in a \caz{similarity}{diffusion} matrix $F$. $F$ is used to compute the similarity matrix $E$ that \caz{accounts also for similarities in node scores as}{models exchanged heat as}
\begin{equation*} 
E = F \operatorname{diag}(w(V)), 
\end{equation*}
where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal. \caz{}{For any two nodes/genes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.}  To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$. Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores. We \caz{scored the nodes}{assigned initial node scores} as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be. In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}. HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\). Both parameters are set automatically by the algorithm, \caz{and are robust}{which is robust to their values} \cite{leiserson_pan-cancer_2015}. HotNet2 is implemented in Python \cite{hotnet2}.

@@

% \begin{equation}
% \label{eq:scones}
% \underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
% \end{equation}

\cazcom{}{Your equation for SConES was wrong.
Use either of the following (I'd use the one that does not involve the Laplacian, then you do not need to introduce the Laplacian in the Mathematical notations.
One way to look at it is to start from the bottom one, where the connectivity term increases with each edge that connects a selected node to a non-selected node, and as you're maximizing the whole thing, you need a minus in top of it. Then 
\[
\sum_{u \in V_S} \sum_{v \not\in V_S} W_{uv} = \sum_{u \in V_S} \left(\sum_{v \in V} W_{uv} - \sum_{v \in V_S} W_{uv} \right) = \sum_{u \in V_S} \left(D_{uu} - \sum_{v \in V_S} W_{uv} \right) 
= \sum_{u \in V_S} \sum_{v \in V_S} \left(D_{uv} - W_{uv} \right) \]
The first equality is true because the set of all nodes is the union of the set of nodes of $S$ and the set of nodes not in $S$; the second by definition of the degree; and the third because the degree matrix is diagonal and $D_{uv}=0$ for $u \neq v$. Finally $L_{uv} = D_{uv} - W_{uv}$.
}

\begin{equation}
\label{eq:scones}
\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} {\color{red}-} \underbrace{\lambda \sum_{v \in V_S} \sum_{u {\color{red} \in} V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
\end{equation}

or
